---
title: "STROKE PREDICTION & PROGNOSIS - LOGISTIC"
author: "Enock Bereka"
date: "2025-05-17"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load necessary libraries
```{r message=FALSE, warning=FALSE}
library(Boruta)
library(tidyverse)
library(caret)
library(pROC)
library(caTools)
```

## Setting working directory
```{r message=FALSE, warning=FALSE}
setwd("C:/Users/DELL/Desktop/Stroke/")
```

## Load the dataset
```{r message=FALSE, warning=FALSE}
data = read_csv("Stroke.csv")
```

## Assess the structure of the dataset
```{r message=FALSE, warning=FALSE}
glimpse(data)
```

## Checking for missing values
```{r message=FALSE, warning=FALSE}
anyNA(data)
```

## Checking for duplicates
```{r message=FALSE, warning=FALSE}
sum(duplicated(data))
```

## Data Preprocessing
## Label encoding
```{r message=FALSE, warning=FALSE}
categorical_vars = c("gender", "ever_married", "Residence_type","work_type", "smoking_status")
data = data %>% mutate(across(all_of(categorical_vars), ~as.numeric(as.factor(.))-1))
```

## Convert categorical variables to factors
```{r message=FALSE, warning=FALSE}
data[categorical_vars] = lapply(data[categorical_vars], as.factor)
```

## Convert BMI to numeric
```{r message=FALSE, warning=FALSE}
data$bmi = as.numeric(data$bmi)
```

## Remove NAs in BMI
```{r message=FALSE, warning=FALSE}
data = drop_na(data)
```

## Remove unnecessary columns
```{r message=FALSE, warning=FALSE}
data$id = NULL
```

## Convert stroke to binary
```{r message=FALSE, warning=FALSE}
data$stroke = ifelse(data$stroke == 1, "Yes", "No")
data$stroke = as.factor(data$stroke)
```

## Checking the distribution of the study outcome
```{r message=FALSE, warning=FALSE}
table(data$stroke)
```

## Handling Class imbalance
## Apply downsampling of the majority class
```{r message=FALSE, warning=FALSE}
data = downSample(x = data[, -11], y = data$stroke)
table(data$Class)
```

## Feature importance analysis
```{r message=FALSE, warning=FALSE}
boruta_results = Boruta(Class~., data)
print(boruta_results)
```

## Check only the confirmed attributes
```{r message=FALSE, warning=FALSE}
getSelectedAttributes(boruta_results, withTentative = T)
```

## Remove the rejected attributes
```{r message=FALSE, warning=FALSE}
data$gender = NULL
data$Residence_type = NULL
```

## Splitting the data into training and testing sets
```{r message=FALSE, warning=FALSE}
index = sample.split(data$Class, SplitRatio = 0.8)
trainData = subset(data, index == "TRUE")
testData = subset(data, index == "FALSE")
table(trainData$Class)
table(testData$Class)
```

## Prepare a training scheme
```{r message=FALSE, warning=FALSE}
fitControl = trainControl(method = "repeatedcv",
                          number = 10,
                          repeats = 4,
                          classProbs = TRUE,
                          summaryFunction = twoClassSummary)
```

## Train the Logistic Regression Model
```{r message=FALSE, warning=FALSE} 
model = train(Class~.,
              trainData,
              method = "glm",
              metric = "ROC",
              preProces = c("center", "scale"),
              trControl = fitControl)
```

## Make predictions
```{r message=FALSE, warning=FALSE}
model_pred = predict(model, testData)
cm = confusionMatrix(model_pred, testData$Class, positive = "Yes")
cm
```

## ROC curve 
```{r message=FALSE, warning=FALSE}
pred =  predict(model, testData, type = "prob")
roc(testData$Class , pred$Yes)
```

